---
title: "updated-rank-ethnicity"
author: "George WIlloughby"
date: "30/07/2021"
output: html_document
---

# Loading the data 

New data was released (July 28) detailing the specific breakdown of ranks and roles in the 43 police forces in England and Wales. The aim of this markdown is to do some simple cleaning and prepare a dataset ready for filtering in Google Sheets. The end point is a dataset where we can just change the force name and ethnicity.

```{r}
ethnicity_rank_2021 <- read.csv("updated-police-workforce-ethnicity.csv")

ethnicity_rank_2021
```


# Changing the column names

The first thing to do is change the column header names to make them easier to understand.

```{r}
#Load library
library(dplyr)

#Check the column names
colnames(ethnicity_rank_2021)

#Rename columns
ethnicity_rank_2021 <- ethnicity_rank_2021 %>%
  rename(
    as_at_31_march = As.at.31.March...,
    geo_code = Geo.code,
    force = Force.Name,
    region = Region,
    sex = Gender,
    rank = Rank.description,
    role = Worker.type,
    ethnicity = Ethnicity,
    ethnic_group = Ethnic.group,
    total_headcount = Total..headcount.,
    total_full_time_equivalent = Total..FTE.
  )
#Check changes
colnames(ethnicity_rank_2021)
```

# Cleaning the columns

The next part of the cleaning process is to select the columns we want to use in our dataset. There are some we don't need as the focus is **police officers** only. This first step is only the start as we will be pulling specific ranks and roles later on in the script.

```{r}
#Selecting columns
ethnicity_rank_2021 <- ethnicity_rank_2021[c(1:8,10)]

ethnicity_rank_2021
```


# Removing unwanted ranks and roles

The purpose of this analysis is to look at police in officer roles. This means we need to remove any administrative roles and also PCSO positions. The officer positions are as follows: National Police Chiefs Council *(NPCC)*,	Chief Superintendent, 	Superintendent, 	Chief Inspector, 	Inspector	Sergeant 	Constable. *NPCC* is also Chief Officer which is what the rank is referred to in the raw data.

You will see then when you drop the ranks, it will also remove the part-time roles in the force. This is what we need as in the **role** columns, we only want **Police Officer**

```{r}
#Look at the officer roles in our dataset
unique(ethnicity_rank_2021$rank)

library(dplyr)
#Dropping unwanted ranks
ethnicity_rank_2021 <- subset(ethnicity_rank_2021, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer" & rank!= "Special Constable")

#Dropping unwanted roles
ethnicity_rank_2021 <- subset(ethnicity_rank_2021, role!= "Police Support Volunteer")

#Check the results
unique(ethnicity_rank_2021$rank)
```

# Selectng 2021

To help with analysis, one last optional code we are going to run is to remove all the years we don't want. The aim is to end up with just the rank figures for 2021, but, having data for all the years is useful so on this occasion, we will save a nee variable.

```{r}
#New variable for 2021 figures
latest_year_ethnicity_rank_2021 <- ethnicity_rank_2021[ethnicity_rank_2021$as_at_31_march >= "2021" & ethnicity_rank_2021$as_at_31_march <= "2021",]

latest_year_ethnicity_rank_2021
```




# Selecting an indvidual force

In terms of cleaning, we have finished in R. We can now export the dataset and continue analysing the figures in a spreadsheet tool. 

```{r}
#Export file
write.csv(ethnicity_rank_2021, "updated-rank-ethnicity-figures-2021.csv")
#Latest year file
write.csv(latest_year_ethnicity_rank_2021, "latest-year-ethnicity-rank-forces-2021.csv")
```






