---
title: "forces-by-ethnicity"
author: "George WIlloughby"
date: "10/06/2021"
output: html_document
---

# Loading the data

The first thing we are going to do is load the data tables provided by the home office. To access these, you can follow this [link](https://www.gov.uk/government/statistics/police-workforce-open-data-tables#history). We are looking for the *ethnicity open data tables* which breaks down each police force area by year, gender, rank, role and ethnicity. We will call our new sheet **ethnicity_data_tables**

```{r}
#Loading the data 
ethnicity_data_tables <- read.csv("data-table-forces-ethnicity.csv", header=T, na.strings=c("","NA"))

ethnicity_data_tables
```


# Changing the column names

Once you have viewed the data, you will probably see that it needs a little cleaning. Firstly, we can change the column headers to make them a bit more readable. After that, we can remove a few columns that we don't need.

```{r}
library(dplyr)

ethnicity_data_tables <- ethnicity_data_tables %>%
  rename(
    as_at_31_march = As.at.31.March,
    geo_code = Geo.code,
    force = Force.Name,
    region = Region,
    sex = Sex,
    rank = Rank.description,
    role = Worker.type,
    ethnicity = Ethnicity,
    ethnic_group = Ethnic.group,
    total_headcount = Total..headcount.,
    total_full_time_equivalent = Total..FTE.
  )

#Selecting the columns we want to keep by index number
ethnicity_data_tables <- ethnicity_data_tables[c(1:11)]

ethnicity_data_tables
```

# Further cleaning

Some police forces might not have a certain ethnicity for a specific officer role. In the data, it is currently shown as an '-'. Let's replace it with a zero.


```{r}
#View the years in the dataframe
unique(ethnicity_data_tables$as_at_31_march)

#Make cells with a '-' an N/A
ethnicity_data_tables$total_headcount[ethnicity_data_tables$total_headcount == "-"] <- NA

ethnicity_data_tables
```

# A closer look at the latest figures

Now that the data has been cleaned up a bit, let's start to take a closer look at the most recent year. To look at just the 2020 figures, we are going to create a new variable so name it accordingly. In this instance, let's go with *ethnicity_data_tables_2020*. 


```{r}
#New variable for 2020 figures
ethnicity_data_tables_2020 <- ethnicity_data_tables[ethnicity_data_tables$as_at_31_march >= "2020" & ethnicity_data_tables$as_at_31_march <= "2020",]

ethnicity_data_tables_2020
```

# Selecting individual forces 

To make analysing each force easier, we can select one force at a time. To do this, wan use the filter tool. The main reason why we want to do this is because we will need to calculate some text values using dplyr() later on and we don't want any potential figures that includes more than one force.

```{r}
library(dplyr)
#View the police force names
unique(ethnicity_data_tables_2020$force)

#Creating a new variable for a specific force
cheshire2020 <- filter(ethnicity_data_tables_2020, force == "Cheshire")

#View the individual force
cheshire2020
```


# Removing unwanted police ranks

The purpose of this analysis is to look at police in officer roles. This means we need to remove any administrative roles and also PCSO positions. Based off FOI request, the officer positions are as follows: National Police Chiefs Council,	Chief Superintendent, 	Superintendent, 	Chief Inspector, 	Inspector	Sergeant 	Constable.

```{r}
#Look at the officer roles in our dataset
unique(cheshire2020$rank)

library(dplyr)
#Dropping unwanted ranks
cheshire2020 <- subset(cheshire2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(cheshire2020$rank)
```
```{r}
#Repeating the same process for a different force

#Creating a new variable for a specific force
gloucester2020 <- filter(ethnicity_data_tables_2020, force == "Gloucestershire")

#View the individual force
gloucester2020

library(dplyr)
#Dropping unwanted ranks
gloucester2020 <- subset(gloucester2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(gloucester2020$rank)
```

```{r}
#Repeating the same process for a different force

#Creating a new variable for a specific force
warwickshire2020 <- filter(ethnicity_data_tables_2020, force == "Warwickshire")

#View the individual force
warwickshire2020

library(dplyr)
#Dropping unwanted ranks
warwickshire2020 <- subset(warwickshire2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(warwickshire2020$rank)
```

```{r}
#Repeating the same process for a different force

#Creating a new variable for a specific force
lancashire2020 <- filter(ethnicity_data_tables_2020, force == "Lancashire")

#View the individual force
lancashire2020

library(dplyr)
#Dropping unwanted ranks
lancashire2020 <- subset(lancashire2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(lancashire2020$rank)
```

```{r}
#Creating a new variable for a specific force
leicestershire2020 <- filter(ethnicity_data_tables_2020, force == "Leicestershire")

#View the individual force
leicestershire2020

library(dplyr)
#Dropping unwanted ranks
leicestershire2020 <- subset(leicestershire2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(leicestershire2020$rank)
```

```{r}
#Creating a new variable for a specific force
humberside2020 <- filter(ethnicity_data_tables_2020, force == "Humberside")

#View the individual force
humberside2020

library(dplyr)
#Dropping unwanted ranks
humberside2020 <- subset(humberside2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(humberside2020$rank)
```

```{r}
#Creating a new variable for a specific force
city_of_london_2020 <- filter(ethnicity_data_tables_2020, force == "London, City of")

#View the individual force
city_of_london_2020

library(dplyr)
#Dropping unwanted ranks
city_of_london_2020 <- subset(city_of_london_2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(city_of_london_2020$rank)
```

```{r}
#Creating a new variable for a specific force
cumbria2020 <- filter(ethnicity_data_tables_2020, force == "Cumbria")

#View the individual force
cumbria2020

library(dplyr)
#Dropping unwanted ranks
cumbria2020 <- subset(cumbria2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(cumbria2020$rank)
```

```{r}
#Creating a new variable for a specific force
west_midlands2020 <- filter(ethnicity_data_tables_2020, force == "West Midlands")

#View the individual force
west_midlands2020

library(dplyr)
#Dropping unwanted ranks
west_midlands2020 <- subset(west_midlands2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(west_midlands2020$rank)
```


```{r}
#Creating a new variable for a specific force
hertfordshire2020 <- filter(ethnicity_data_tables_2020, force == "Hertfordshire")

#View the individual force
hertfordshire2020

library(dplyr)
#Dropping unwanted ranks
hertfordshire2020 <- subset(hertfordshire2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(hertfordshire2020$rank)
```

```{r}
#Creating a new variable for a specific force
bedfordshire2020 <- filter(ethnicity_data_tables_2020, force == "Bedfordshire")

#View the individual force
bedfordshire2020

library(dplyr)
#Dropping unwanted ranks
bedfordshire2020 <- subset(bedfordshire2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(bedfordshire2020$rank)
```

```{r}
#Creating a new variable for a specific force
south_yorkshire2020 <- filter(ethnicity_data_tables_2020, force == "South Yorkshire")

#View the individual force
south_yorkshire2020

library(dplyr)
#Dropping unwanted ranks
south_yorkshire2020 <- subset(south_yorkshire2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(south_yorkshire2020$rank)
```
```{r}
#Creating a new variable for a specific force
dyfed_powys2020 <- filter(ethnicity_data_tables_2020, force == "Dyfed-Powys")

#View the individual force
dyfed_powys2020

library(dplyr)
#Dropping unwanted ranks
dyfed_powys2020 <- subset(dyfed_powys2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(dyfed_powys2020$rank)
```

```{r}
#Creating a new variable for a specific force
lincolnshire2020 <- filter(ethnicity_data_tables_2020, force == "Lincolnshire")

#View the individual force
lincolnshire2020

library(dplyr)
#Dropping unwanted ranks
lincolnshire2020 <- subset(lincolnshire2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(lincolnshire2020$rank)
```

```{r}
#Creating a new variable for a specific force
merseyside2020 <- filter(ethnicity_data_tables_2020, force == "Merseyside")

#View the individual force
merseyside2020

library(dplyr)
#Dropping unwanted ranks
merseyside2020 <- subset(merseyside2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(merseyside2020$rank)
```


```{r}
#Creating a new variable for a specific force
kent2020 <- filter(ethnicity_data_tables_2020, force == "Kent")

#View the individual force
kent2020

library(dplyr)
#Dropping unwanted ranks
kent2020 <- subset(kent2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(kent2020$rank)
```
```{r}
#Creating a new variable for a specific force
staffordshire2020 <- filter(ethnicity_data_tables_2020, force == "Staffordshire")

#View the individual force
staffordshire2020

library(dplyr)
#Dropping unwanted ranks
staffordshire2020 <- subset(staffordshire2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(staffordshire2020$rank)
```
```{r}
#Creating a new variable for a specific force
west_mercia2020 <- filter(ethnicity_data_tables_2020, force == "West Mercia")

#View the individual force
west_mercia2020

library(dplyr)
#Dropping unwanted ranks
west_mercia2020 <- subset(west_mercia2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(west_mercia2020$rank)
```
```{r}
#Creating a new variable for a specific force
sussex2020 <- filter(ethnicity_data_tables_2020, force == "Sussex")

#View the individual force
sussex2020

library(dplyr)
#Dropping unwanted ranks
sussex2020 <- subset(sussex2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(sussex2020$rank)
```

```{r}
#Creating a new variable for a specific force
northamptonshire2020 <- filter(ethnicity_data_tables_2020, force == "Northamptonshire")

#View the individual force
northamptonshire2020

library(dplyr)
#Dropping unwanted ranks
northamptonshire2020 <- subset(northamptonshire2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(northamptonshire2020$rank)
```
```{r}
#Creating a new variable for a specific force
gwent2020 <- filter(ethnicity_data_tables_2020, force == "Gwent")

#View the individual force
gwent2020

library(dplyr)
#Dropping unwanted ranks
gwent2020 <- subset(gwent2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(gwent2020$rank)
```
```{r}
#Creating a new variable for a specific force
essex2020 <- filter(ethnicity_data_tables_2020, force == "Essex")

#View the individual force
essex2020

library(dplyr)
#Dropping unwanted ranks
essex2020 <- subset(essex2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(essex2020$rank)
```
```{r}
#Creating a new variable for a specific force
cleveland2020 <- filter(ethnicity_data_tables_2020, force == "Cleveland")

#View the individual force
cleveland2020

library(dplyr)
#Dropping unwanted ranks
cleveland2020 <- subset(cleveland2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(cleveland2020$rank)
```

```{r}
#Creating a new variable for a specific force
cambridgeshire2020 <- filter(ethnicity_data_tables_2020, force == "Cambridgeshire")

#View the individual force
cambridgeshire2020

library(dplyr)
#Dropping unwanted ranks
cambridgeshire2020 <- subset(cambridgeshire2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(cambridgeshire2020$rank)
```
```{r}
#Creating a new variable for a specific force
norfolk2020 <- filter(ethnicity_data_tables_2020, force == "Norfolk")

#View the individual force
norfolk2020

library(dplyr)
#Dropping unwanted ranks
norfolk2020 <- subset(norfolk2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(norfolk2020$rank)
```

```{r}
#Creating a new variable for a specific force
thamesvalley2020 <- filter(ethnicity_data_tables_2020, force == "Thames Valley")

#View the individual force
thamesvalley2020

library(dplyr)
#Dropping unwanted ranks
thamesvalley2020 <- subset(thamesvalley2020, rank!="Designated Detention Officer (S.38)" & rank!="Community Support Officer" & rank!="Designated Detention Officer (S.39)" & rank!= "Designated Escort Officer (S.38)" & rank!="Designated Escort Officer (S.39)" & rank!= "Designated Investigation Officer (S.38)" & rank!="Police Staff" & rank!="Police support volunteer")

#Check the results
unique(thamesvalley2020$rank)
```




#Exporting the data

In terms of analysis, we are at a stage now where we can export. We have the most up to date year, we have a script that narrows down to an individual force and some simple filtering in excel is the only thing required post export. To alter the script so that you can look at othet forces, simply change the force filter in the **Selecting individual forces** section to the one the one you want to look at.

```{r}
#Exporting as a CSV file
write.csv(cheshire2020, "cheshire-police-force-breakdown.csv")
write.csv(gloucester2020, "gloucester-police-force-breakdown.csv")
write.csv(warwickshire2020, "warwickshire-police-force-breakdown.csv")
write.csv(lancashire2020,"lancashire-police-force-breakdown.csv")
write.csv(leicestershire2020,"leicestershire-police-force-breakdown.csv")
write.csv(humberside2020, "humberside-police-fore-breakdown.csv")
write.csv(city_of_london_2020, "city-of-london-police-force-breakdown.csv")
write.csv(cumbria2020, "cumbria-police-force-breakdown.csv")
write.csv(west_midlands2020, "west-midlands-police-force-breakdown.csv")
write.csv(hertfordshire2020, "hertfordshire-police-force-breakdown.csv")
write.csv(bedfordshire2020, "bedfordshire-police-force-breakdown.csv")
write.csv(south_yorkshire2020, "south-yorkshire-police-force-breakdown.csv")
write.csv(dyfed_powys2020, "dyfed-powys-police-force-breakdown.csv")
write.csv(lincolnshire2020, "lincolnshire-police-force-breakdown.csv")
write.csv(merseyside2020, "merseyside-police-force-breakdown.csv")
write.csv(kent2020, "kent-police-force-breakdown.csv")
write.csv(staffordshire2020, "staffordshire-police-force-breakdown.csv")
write.csv(west_mercia2020, "west-mercia-police-force-breakdown.csv")
write.csv(sussex2020, "sussex-police-force-breakdown.csv")
write.csv(northamptonshire2020, "northamptonshire-police-force-breakdown.csv")
write.csv(gwent2020, "gwent-police-force-breakdown.csv")
write.csv(essex2020, "essex-police-force-breakdown.csv")
write.csv(cleveland2020, "cleveland-police-force-breakdown.csv")
write.csv(cambridgeshire2020, "cambridgeshire-police-force-breakdown.csv")
write.csv(norfolk2020, "norfolk-police-force-breakdown.csv")
write.csv(thamesvalley2020, "thames-valley-police-force-breakdown.csv")
```



cambridgeshire2020



